---
title: "Severe Weather Events of Greatest Impact"
author: "jmtaysom"
date: "April 23, 2015"
output: html_document
---
## Severe Weather Events of Greatest Impact
Synopsis: This project examines NOAA's storm database in an attempt to analyze
the storms and identify those with the greatest impact. There are two categories
that will be the main focus: greatest property damage and most harmful to human
health. As seen in the results Tornadoes harm the most people while flooding and 
huricanes do the most property damage.

## Data Processing

Load the appropriate libraries.
```{r}
library(ggplot2)
library(reshape2)
library(plyr)
```

Download and open the storm dataset.
```{r, cache=TRUE}
setwd("~/datasciencecoursera/ReproducibleResearch")
url = 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2'
zip = 'Storm_Data.csv.bz2'
#download.file(url, zip, method='curl')
rm(url)
storm <- read.csv(bzfile(zip))
rm(zip)
```

Calculate the total damagein billions of dollars by converting the character
descriptions of the magnitude of damage into numeric and then applying it to the
value of the damage.
```{r}
storm$PROPDMGEXP <-revalue(storm$PROPDMGEXP, c("K"="1000", "M"="1000000", "B"="1000000000"))
storm$PROPDMGEXP <- sub("^$", "1", storm$PROPDMGEXP)
storm$PROPDMG <- storm$PROPDMG * as.numeric(storm$PROPDMGEXP) / 1000000000

```

Sumarize the injuries and fatalities by event type and display the most damaging
events. In the top 6 there are 5 different event types with tornadoes taking
two of the spots since more people died from tornadoes than were injured by the 
next most serious weather type.
```{r}
human.damage <- ddply(storm, "EVTYPE", summarize, injuries = sum(INJURIES), fatalities = sum(FATALITIES))
hd.m <- melt(human.damage, id.vars='EVTYPE')
hd.m <- hd.m[complete.cases(hd.m),]
top.hd <- tail(arrange(hd.m, value), n=6)
top.hd
```

Sumarize the total damage by event type and select the top three which happen to
do an order of magnitude more damage than the next few.
```{r}
prop.damage <- ddply(storm, "EVTYPE", summarize, propdmg = sum(PROPDMG))
prop.damage <- prop.damage[complete.cases(prop.damage),]
top.prop <-tail(arrange(prop.damage, propdmg), n=3)
```
##Results

Tornadoes are by far the greatest threat to human health. Tornado fatalities 
rank 5th in the most people effected by weather and there are more than 10 times
the number of injuries from tornadoes as any other other weather event.
```{r}
ggplot(top.hd, aes(reorder(x=EVTYPE, -value), y=value,fill=variable)) + geom_bar(stat='identity') + labs(x="Storm Type", y="People injured and killed")
```

Floods and huricanes are the weather events that do the most property damage. 
The top three property damaging weather events do an order of magnitude more 
damage than any of the other events.
```{r}
ggplot(top.prop, aes(x=EVTYPE, y=propdmg)) + geom_bar(stat='identity') + labs(x="Storm Type", y="Property Damage in Billions of Dollars")
```


