---
title: "Statistical Inference Project: Part 1"
author: "jmtaysom"
date: "05/08/2015"
output:
  pdf_document: default
---
```{r load.library, echo=FALSE, results='hide'}
library(ggplot2)
library(grid)
library(gridExtra)
```
```{r load.data, echo=FALSE, results='hide'}
set.seed(0)
lambda <- 0.2
observations <- 40
trials <- 1000

expn = NULL
for (i in 1:trials) expn = c(expn, mean(rexp(observations, lambda)))

exsigma <- (1/lambda)/sqrt(observations)
exmean <- (1/lambda)
acsigma <- sd(expn)
acmean <- mean(expn)

norm <- rnorm(trials,mean=1/lambda, sd = (1/lambda)/sqrt(observations))
```

# Executive summary
In this report we will explore the Central Limit Therom and how it applys to multiple draws of a sample from a population. Specifically we will look at how `r observations` means of draws of sample of size `r observations` random exponential numbers with a lambda of `r lambda` compares to `r trials` random normal observations. The random normal numbers will have a mean and standard deviation equivilant to the theoretical mean and standard deviation that we would expect for the exponential numbers. The mean and standard deviation of the population of eponential numbers are both equal to $\frac{1}{\lambda}$ or in ths case equal to `r 1/lambda`. The theoretical standard deviation of the sample will be equal to $\frac{\sigma}{\sqrt{`r observations`}}$ on in this case `r (1/lambda)/sqrt(observations)`. The actuall mean of the `r trials` means of `r observations` exponential numbers is `r mean(expn)` and the standard deviation is `r sd(expn)` show that the observed values are representitive of the population according to the Central Limit Therom.

# Introduction 
In this experiment we will be comparing the distribution of the of `r trials`  means of samples of `r observations` exponential numbers to their expected distribution through the use of the central limit therom. The poultaion of exponential numbers will have a mean and standard deviation of $\frac{1}{\lambda}$ or $\frac{1}{0.2}=`r 1/lambda`$. The standard deviation of the sample will be equal to $\frac{\sigma}{\sqrt{`r observations`}}$ on in this case `r (1/lambda)/sqrt(observations)`.

# Calculations
## Comparing to expected values
Upon sampling the exponential numbers we can start comparing them to the expected values. The expected mean of the samples is `r exmean` while the observed mean is `r acmean` as seen in Figure 1 in the appendix. The expected variance is `r exsigma^2` and the observed variance is `r acsigma^2`. However with the sample size being `r observations` the variance is much smaller than the variance of the population. The variance of the population would be $(\frac{1}{\lambda})^2$ or `r (1/lambda)^2` which is  greater than the observed variance as seen in figure 2 in the appendix.  With the observed mean and variance so close to the expected mean and standard deviation we can see that the central limit therom governs this dataset.

## Comparing to normal data
While the distibution of exponential numbers is not normal the distribution of the mean of the samples should be normal. This can be tested by comparing the distibution of the samples to the distribution of the same number of random normal numbers that have a mean and standard deviation equal to that of the samples of exponential numbers. The result as seen in figure 3 show that distibution of the means of the samples of exponential numbers are distributed normally.

# Appendix

## Figures

### Figure 1
Histogram of the sampled expoential numbers with the expected mean and actual mean
```{r graph_expn}
xdf <- data.frame(expn)
x <- ggplot(xdf,aes(x=expn)) + geom_histogram(aes(fill = ..count..), binwidth=.25) + xlim(2,8) +ylim(0,150)
x + geom_vline(xintercept = mean(expn), color='red') + geom_vline(xintercept = 1/lambda, color='black') + ggtitle('Expected Mean (black) vs. Actual Mean (red)' )
```

### Figure 2
The variance of the sample is smaller than the variance of the poplutation
```{r graph_distributions}
norm <-rnorm(1000,mean=5, sd = 5)
ndf <- data.frame(norm)
ndf$distribution <- 'population'
names(ndf)[names(ndf)=="norm"] <- "value"
xdf$distribution <- 'sample'
names(xdf)[names(xdf)=="expn"] <- "value"
combined <- rbind(xdf,ndf)

ggplot(combined, aes(x=value, colour=distribution)) + geom_density() +ggtitle('Sample Variance vs. Population Variance')
```

### Figure 3
Comparing the distribution of the samples to a normal distribution
```{r graph_normal}
norm <-rnorm(1000,mean=5, sd = 5/sqrt(40))
ndf <- data.frame(norm)
n <- ggplot(ndf, aes(x=norm)) + geom_histogram(aes(fill = ..count..), binwidth=.25) + xlim(2,8) +ylim(0,150)

grid.arrange(x,n, ncol=2, main='Exponential distribution vs. normal distribution')
```


## Loading Data
Here we set the seed to verify the reproducibility of the results. Set the variables according to the experiment. And create two samples; one for `r trials` of the mean of `r observations` exponential numbers, and one for `r trials` of normal numbers that have a mean and sd equal to the exponential numbers.
```{r show.load.data, ref.label='load.data'}
```